---
title: "Likert scale"
author: "Daniela Palleschi"
date: "`r Sys.Date()`"
format: 
  html:
    toc: true
    execute:
      echo: true
      eval: true
      warning: false
      message: false
editor_options: 
  chunk_output_type: console
---

This script loads in example Likert-scale data and produces plots and summary tables. The plots and tables are then saved externally.

# Packages

Install necessary packages if not done already.

```{r}
#| code-fold: true
#| eval: false
install.packages("pacman")
```

Load in packages

```{r}
#| code-fold: true
pacman::p_load(tidyverse,
               here,
               knitr,
               kableExtra,
               magick)
```


# read in data

```{r}
df_toydata <- read_csv(here::here("data", "df_toydata.csv")) %>% 
  mutate(condition = paste(heritage,instruments, sep="_")) %>% 
  mutate(tempo_centred = tempo_rating - 4,
         familiarity_centred = familiarity_rating - 4,
         formality_centred = formality_rating - 4,
         formality_diff_centred = formality_differences - 4
         )
```

```{r}
df_toydata %>% summary() %>% kbl() %>% kable_styling()
```

# Plots

## All scales by language

Create preferred facet labels.

```{r}
facet_labels <-
  c(
    "tempo_centred" = "Tempo",
    "familiarity_centred" =  "Familiarity",
    "formality_centred" = "Formality",
    "formality_diff_centred" = "Formality Difference"
    
  )
```

Generate plot.

```{r}
#| code-fold: true

fig_scales_heritage <-
  df_toydata %>%
  mutate(tempo_centred = factor(tempo_centred,
                                levels = c("-3", "-2", "-1", "0", "1", "2", "3"))) %>%
  mutate(familiarity_centred = factor(familiarity_centred, levels = c("-3", "-2", "-1", "0", "1", "2", "3")
  )) %>%
  mutate(formality_centred = factor(formality_centred, levels = c("-3", "-2", "-1", "0", "1", "2", "3"))) %>%
  mutate(formality_diff_centred = factor(formality_diff_centred, levels = c("-3", "-2", "-1", "0", "1", "2", "3")
  )) %>% 
  pivot_longer(
    cols = c(
      tempo_centred,
      familiarity_centred,
      formality_centred,
      formality_diff_centred
    ),
    names_to = "scale",
    values_to = "rating"
  ) %>%
  mutate(rating = reorder(rating, desc(rating))) %>%
  # aes(x = reorder(the_factor, desc(the_factor))
  ggplot(aes(y = heritage, fill = rating)) +
  labs(title = "Ratings per scale",
       x = "Count",
       y = "Heritage") +
  facet_grid(scale ~ ., labeller = as_labeller(facet_labels)) +
  geom_bar() +
  geom_vline(xintercept = 50) +
  scale_fill_manual(values = c(
    "darkgreen",
    "chartreuse3",
    "lightgreen",
    "grey",
    "coral",
    "red3",
    "red4"
  )) +
  theme_bw() +
  theme(legend.position = "bottom") + guides(fill = guide_legend(nrow = 1, reverse = TRUE))
```

```{r}
fig_scales_heritage
```

Save plot.

```{r}
ggsave(here("plots","barplot_4scales_heritage.jpeg"),
       fig_scales_heritage,
       scale = 1.3,
  width = 210,
  height = 125,
  units = c("mm"))
```

## Plot per scale

### Familiarity

```{r}
#| code-fold: true
fig_familiarity <-
  df_toydata %>%
  mutate(familiarity_centred = factor(
    familiarity_centred,
    levels = c("-3", "-2", "-1", "0", "1", "2", "3")
  )) %>%
  mutate(familiarity_centred = reorder(familiarity_centred, desc(familiarity_centred))) %>%
  # aes(x = reorder(the_factor, desc(the_factor))
  ggplot(aes(y = heritage, fill = familiarity_centred)) +
  labs(title = "Familiarity ratings",
       x = "Count",
       y = "Heritage",
       fill = "Rating") +
  # facet_grid(scale ~ ., labeller = as_labeller(facet_labels)) +
  geom_bar() +
  geom_vline(xintercept = 50) +
  scale_fill_manual(values = c(
    "darkgreen",
    "chartreuse3",
    "lightgreen",
    "grey",
    "coral",
    "red3",
    "red4"
  )) +
  theme_bw() +
  theme(legend.position = "bottom") + guides(fill = guide_legend(nrow = 1, reverse = TRUE))
```

```{r}
fig_familiarity
```

Save plot.

```{r}
ggsave(here("plots","barplot_familiarty.jpeg"),
       fig_familiarity,
       scale = 1.3,
  width = 210,
  height = 80,
  units = c("mm"))
```


### Formality

```{r}
#| code-fold: true
fig_formality <-
  df_toydata %>%
  mutate(formality_centred = factor(
    formality_centred,
    levels = c("-3", "-2", "-1", "0", "1", "2", "3")
  )) %>%
  mutate(formality_centred = reorder(formality_centred, desc(formality_centred))) %>%
  # aes(x = reorder(the_factor, desc(the_factor))
  ggplot(aes(y = heritage, fill = formality_centred)) +
  labs(title = "Formality ratings",
       x = "Count",
       y = "Heritage",
       fill = "Rating") +
  # facet_grid(scale ~ ., labeller = as_labeller(facet_labels)) +
  geom_bar() +
  geom_vline(xintercept = 50) +
  scale_fill_manual(values = c(
    "darkgreen",
    "chartreuse3",
    "lightgreen",
    "grey",
    "coral",
    "red3",
    "red4"
  )) +
  theme_bw() +
  theme(legend.position = "bottom") + guides(fill = guide_legend(nrow = 1, reverse = TRUE))
```

```{r}
fig_formality
```

Save plot.

```{r}
ggsave(here("plots","barplot_familiarty.jpeg"),
       fig_formality,
       scale = 1.3,
  width = 210,
  height = 80,
  units = c("mm"))
```

### Tempo

```{r}
#| code-fold: true
fig_tempo <-
  df_toydata %>%
  mutate(tempo_centred = factor(
    tempo_centred,
    levels = c("-3", "-2", "-1", "0", "1", "2", "3")
  )) %>%
  mutate(tempo_centred = reorder(tempo_centred, desc(tempo_centred))) %>%
  # aes(x = reorder(the_factor, desc(the_factor))
  ggplot(aes(y = heritage, fill = tempo_centred)) +
  labs(title = "Tempo ratings",
       x = "Count",
       y = "Heritage",
       fill = "Rating") +
  # facet_grid(scale ~ ., labeller = as_labeller(facet_labels)) +
  geom_bar() +
  geom_vline(xintercept = 50) +
  scale_fill_manual(values = c(
    "darkgreen",
    "chartreuse3",
    "lightgreen",
    "grey",
    "coral",
    "red3",
    "red4"
  )) +
  theme_bw() +
  theme(legend.position = "bottom") + guides(fill = guide_legend(nrow = 1, reverse = TRUE))
```

```{r}
fig_tempo
```

Save plot.

```{r}
ggsave(here("plots","barplot_tempo.jpeg"),
       fig_tempo,
       scale = 1.3,
  width = 210,
  height = 80,
  units = c("mm"))
```

### Formality differences

```{r}
#| code-fold: true
fig_formality_diff <-
  df_toydata %>%
  mutate(formality_diff_centred = factor(
    formality_diff_centred,
    levels = c("-3", "-2", "-1", "0", "1", "2", "3")
  )) %>%
  mutate(formality_diff_centred = reorder(formality_diff_centred, desc(formality_diff_centred))) %>%
  # aes(x = reorder(the_factor, desc(the_factor))
  ggplot(aes(y = heritage, fill = formality_diff_centred)) +
  labs(title = "Formality difference ratings",
       x = "Count",
       y = "Heritage",
       fill = "Rating") +
  # facet_grid(scale ~ ., labeller = as_labeller(facet_labels)) +
  geom_bar() +
  geom_vline(xintercept = 50) +
  scale_fill_manual(values = c(
    "darkgreen",
    "chartreuse3",
    "lightgreen",
    "grey",
    "coral",
    "red3",
    "red4"
  )) +
  theme_bw() +
  theme(legend.position = "bottom") + guides(fill = guide_legend(nrow = 1, reverse = TRUE))
```

```{r}
fig_formality_diff
```

Save plot.

```{r}
ggsave(here("plots","barplot_formality_diff.jpeg"),
       fig_formality_diff,
       scale = 1.3,
  width = 210,
  height = 80,
  units = c("mm"))
```

## Instruments

Create preferred labels for `instruments`.
```{r}
instrument_labels <-
  c(
    "high" = "High",
    "low" =  "Low"
  )
```

### Formal

```{r}
#| code-fold: true
fig_formal_instruments <-
  df_toydata %>%
  filter(heritage == "formal") %>%
  dplyr::mutate(across(
                  c(tempo_centred:formality_diff_centred),
                  ~ factor(., levels = c("-3", "-2", "-1", "0", "1", "2", "3"))
                )) %>% 
  pivot_longer(
    cols = c(tempo_centred:formality_diff_centred
    ),
    names_to = "scale",
    values_to = "rating"
  ) %>%
  mutate(rating = reorder(rating, desc(rating))) %>%
  # aes(x = reorder(the_factor, desc(the_factor))
  ggplot(aes(y = instruments, fill = rating)) +
  labs(title = "Intrument ratings (Formal heritage)",
       x = "Count",
       y = "Instrument",
       fill = "Rating") +
  facet_grid(scale ~ ., labeller=as_labeller(facet_labels)) +
  scale_y_discrete(labels=instrument_labels) +
  geom_bar() +
  geom_vline(xintercept = 25) +
  scale_fill_manual(values = c(
    "darkgreen",
    "chartreuse3",
    "lightgreen",
    "grey",
    "coral",
    "red3",
    "red4"
  )) +
  theme_bw() +
  theme(legend.position = "bottom") + guides(fill = guide_legend(nrow = 1, reverse = TRUE))
```

```{r}
fig_formal_instruments
```

Save plot.

```{r}
ggsave(here("plots", "barplot_formal_instruments.jpeg"),
       fig_formal_instruments,
       scale = 1.3,
  width = 210,
  height = 100,
  units = c("mm"))
```

### Informal

```{r}
#| code-fold: true
fig_informal_instruments <-
  df_toydata %>%
  filter(heritage == "informal") %>%
  dplyr::mutate(across(
                  c(tempo_centred:formality_diff_centred),
                  ~ factor(., levels = c("-3", "-2", "-1", "0", "1", "2", "3"))
                )) %>% 
  pivot_longer(
    cols = c(tempo_centred:formality_diff_centred
    ),
    names_to = "scale",
    values_to = "rating"
  ) %>%
  mutate(rating = reorder(rating, desc(rating))) %>%
  # aes(x = reorder(the_factor, desc(the_factor))
  ggplot(aes(y = instruments, fill = rating)) +
  labs(title = "Intrument ratings (Informal heritage)",
       x = "Count",
       y = "Instrument",
       fill = "Rating") +
  facet_grid(scale ~ ., labeller=as_labeller(facet_labels)) +
  scale_y_discrete(labels=instrument_labels) +
  geom_bar() +
  geom_vline(xintercept = 25) +
  scale_fill_manual(values = c(
    "darkgreen",
    "chartreuse3",
    "lightgreen",
    "grey",
    "coral",
    "red3",
    "red4"
  )) +
  theme_bw() +
  theme(legend.position = "bottom") + guides(fill = guide_legend(nrow = 1, reverse = TRUE))
```

```{r}
fig_informal_instruments
```

Save plot.

```{r}
ggsave(here("plots", "barplot_informal_instruments.jpeg"),
       fig_informal_instruments,
       scale = 1.3,
  width = 210,
  height = 100,
  units = c("mm"))
```

# Summary tables

Create summary tables.

## Language and scale

```{r}
#| code-fold: true
tbl_ratings <-
df_toydata %>% 
  pivot_longer(
    cols = c(
      tempo_centred:formality_diff_centred
    ),
    names_to = "scale",
    values_to = "rating"
  ) %>%
  summarise(mean = mean(rating),
            sd = sd(rating),
            median = median(rating),
            N = n(),
            .by = c(heritage,scale)) %>% 
  knitr::kable(col.names = c("Heritage", "Scale", "M", "SD", "Median", "N"),
               caption = "M = mean, SD = standard deviaton, N = number of ratings") %>% 
  kable_styling()
```

```{r}
tbl_ratings
```

Save table as `.png`.

```{r}
kableExtra::save_kable(tbl_ratings, here("tables", "summary_4scales_heritage.png"))
```

## Instruments

### Formal heritage

```{r}
#| code-fold: true
tbl_formal <-
df_toydata %>% 
  filter(heritage == "formal") %>% 
  pivot_longer(
    cols = c(
      tempo_centred:formality_diff_centred
    ),
    names_to = "scale",
    values_to = "rating"
  ) %>%
  summarise(mean = mean(rating),
            sd = sd(rating),
            median = median(rating),
            N = n(),
            .by = c(heritage, scale, instruments)) %>% 
  knitr::kable(col.names = c("Heritage", "Scale", "Instruments","M", "SD", "Median", "N"),
               caption = "M = mean, SD = standard deviaton, N = number of ratings") %>% 
  kable_styling()
```

```{r}
tbl_formal
```

Save table as `.png`.

```{r}
kableExtra::save_kable(tbl_formal, here("tables", "summary_formal_instruments.png"))
```

### Informal heritage

```{r}
#| code-fold: true
tbl_informal <-
df_toydata %>% 
  filter(heritage == "informal") %>% 
  pivot_longer(
    cols = c(
      tempo_centred:formality_diff_centred
    ),
    names_to = "scale",
    values_to = "rating"
  ) %>%
  summarise(mean = mean(rating),
            sd = sd(rating),
            median = median(rating),
            N = n(),
            .by = c(heritage, scale, instruments)) %>% 
  knitr::kable(col.names = c("Heritage", "Scale", "Instruments","M", "SD", "Median", "N"),
               caption = "M = mean, SD = standard deviaton, N = number of ratings") %>% 
  kable_styling()
```

```{r}
tbl_informal
```

Save table as `.png`.

```{r}
kableExtra::save_kable(tbl_informal, here("tables", "summary_informal_instruments.png"))
```